SOURCE SEPARATION METHODS FOR COMPUTER-ASSISTED ORCHESTRATION

### 2021-06-03 source_separation/paper/paper.tex
Comment out copyrights, update github link
Add index terms

% Copyright notices.
% ------------------
% Select one of the four copyright notices below. Only required for the camera-ready paper submission.
% 
% * For papers in which all authors are employed by the US government:
\copyrightnotice{U.S.\ Government work not protected by U.S.\ copyright}

% * For papers in which all authors are employed by a Crown government (UK, Canada, and Australia):
\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 Crown}

% * For papers in which all authors are employed by the European Union:
\copyrightnotice{978-1-7281-6338-3/21/\$31.00
{\copyright}2021 European Union}

@@

\begin{abstract}
  In this paper, we will study the possibility of adding source separation as a pre-processing step to the computer-assisted orchestration process. We first discuss the motivation of this addition and its potential to increase the quality of orchestrations of multi-layered sounds. Second, we select several state-of-the-art models for both music source separation (separation of instruments) and universal sound separation (separation of arbitrary sounds of different types), and compare their effectiveness for the task of orchestration. We assess which methods best suit the needs of orchestration by applying them on hand picked target sounds, orchestrating the separated outputs, and finally comparing them to the orchestration of the same target without separation. Our experiments show that the quality of orchestrations improves, both qualitatively and quantitatively, indicating that our approach is promising. Finally, we compare unsupervised methods to supervised methods for separation, and comment on the effect of training data selection on performance of supervised methods.
\end{abstract}
%
\begin{keywords}
One, two, three, four, five
\end{keywords}
%
\section{Introduction}\label{sec:introduction}
  Computer-assisted composition is a field that focuses on the creation of computational tools to aid in the musical composition process \cite{FerVic2013, Ari2005}. Within this field, target-based computer-assisted musical orchestration is an example of how musical creativity can be supported by tools such as machine learning. Orchestration has been traditionally taught in an intuitive and non-formalized way, and computer-assisted orchestration seeks to help composers by accelerating certain creative processes. In particular, target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound. Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints~\cite{Maresz2003}.

@@

%  The codebase for this paper can be found at: \url{https://github.com/dzluke/smc2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/smc2021/}.
  The codebase for this paper can be found at: \url{anonymous}, and you can listen to a selection of targets and orchestrations here: \url{anonymous}.

@@

    The procedure we use for testing is as follows:
    \begin{enumerate}
      \item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times.
      \item The full target, without any separation performed, is orchestrated using the entire orchestra. This is called the full target orchestration.
      \item Each separation method is performed on the target, each splitting the target into four sub-targets. 
      \item The full orchestra is randomly split into four equal sized subsets, each containing 7 to 8 instruments.
      \item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously. This creates the \textbf{separated orchestrations}.
      \item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
      \item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration. 
    \end{enumerate}

>>>>

% Copyright notices.
% ------------------
% Select one of the four copyright notices below. Only required for the camera-ready paper submission.
% 
% * For papers in which all authors are employed by the US government:
%\copyrightnotice{U.S.\ Government work not protected by U.S.\ copyright}

% * For papers in which all authors are employed by a Crown government (UK, Canada, and Australia):
%\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 Crown}

% * For papers in which all authors are employed by the European Union:
%\copyrightnotice{978-1-7281-6338-3/21/\$31.00
%{\copyright}2021 European Union}

@@

\begin{abstract}
  In this paper, we will study the possibility of adding source separation as a pre-processing step to the computer-assisted orchestration process. We first discuss the motivation of this addition and its potential to increase the quality of orchestrations of multi-layered sounds. Second, we select several state-of-the-art models for both music source separation (separation of instruments) and universal sound separation (separation of arbitrary sounds of different types), and compare their effectiveness for the task of orchestration. We assess which methods best suit the needs of orchestration by applying them on hand picked target sounds, orchestrating the separated outputs, and finally comparing them to the orchestration of the same target without separation. Our experiments show that the quality of orchestrations improves, both qualitatively and quantitatively, indicating that our approach is promising. Finally, we compare unsupervised methods to supervised methods for separation, and comment on the effect of training data selection on performance of supervised methods.
\end{abstract}
%
\begin{keywords}
computer-assisted orchestration, sound source separation, Orchidea
\end{keywords}
%
\section{Introduction}\label{sec:introduction}
  Computer-assisted composition is a field that focuses on the creation of computational tools to aid in the musical composition process \cite{FerVic2013, Ari2005}. Within this field, target-based computer-assisted musical orchestration is an example of how musical creativity can be supported by tools such as machine learning. Orchestration has been traditionally taught in an intuitive and non-formalized way, and computer-assisted orchestration seeks to help composers by accelerating certain creative processes. In particular, target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound. Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints~\cite{Maresz2003}.

@@

%  The codebase for this paper can be found at: \url{https://github.com/dzluke/MLSP2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/MLSP2021/}.
  The codebase for this paper can be found at: \url{anonymous}, and you can listen to a selection of targets and orchestrations here: \url{anonymous}.

@@

    The procedure we use for testing is as follows:
    \begin{enumerate}
      \item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times.
      \item The target, without any separation performed, is orchestrated using the entire orchestra. This is called the full target orchestration.
      \item Each separation method is performed on the target, each splitting the target into four sub-targets. 
      \item The full orchestra is randomly split into four equal sized subsets, each containing 7 to 8 instruments.
      \item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously. This creates the \textbf{separated orchestrations}.
      \item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
      \item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration. 
    \end{enumerate}
